---
title: "Data Science Project EDA - With Reviewer Notes"
author: "Teagan Zuniga"
date: "2023-11-20"
output: html_document
---

# Research Questions #
    
Q1: Whether age-group affects appraisals of technology
Q2: Whether income-level negatively affects an individuals appraisals of technology (lower = more neg)

# Setup # 

```{r eval = FALSE}
library(renv)
# Using renv restore function to get dependencies 
restore()
```

# Read in data #
```{r}
library(readr)
library(here)

SurveyData <- read_csv(here("data","SurveyData.csv"), show_col_types = FALSE)
```

# Check the packaging #
```{r}
library(skimr)

skim(SurveyData)
```

107 rows, 33 columns *NOTE: THERE ARE 36 COLUMNS, NOT 33*
Data is 100% complete, no concern over missing values 


# Look at the top and the bottom of the data #
```{r}
library(skimr)

if (interactive()) {
  View(head(SurveyData))
  View(tail(SurveyData))
  View(SurveyData)
}
```



## Created Variables and Relevant Overview Plots ##
```{r}
library(ggplot2)

# Age distribution Histogram
ggplot(SurveyData, aes(x = Age)) +
  geom_histogram(
    bins = 20,
    fill = "blue",
    color = "black",
    alpha = 0.7
  ) +
  labs(x = "Age", y = "Count", title = "Distribution of Respondents' Age") +
  theme_minimal()
```

```{r}
library(ggplot2)
library(dplyr)

# Gender distribution pie chart
SurveyData %>%
  count(Gender) %>%
  ggplot(aes(x = "", y = n, fill = Gender)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_fill_brewer(palette = "Pastel1") +
  theme_void() +
  labs(title = "Gender Distribution") +
  geom_text(aes(label = scales::percent(n / sum(n), accuracy = 0.1)),
    position = position_stack(vjust = 0.5)
  )
```


```{r}
# Calculating mean and median age
mean_age <-
  mean(SurveyData$Age, na.rm = TRUE) # na.rm = TRUE ensures that NA values are removed
median_age <- median(SurveyData$Age, na.rm = TRUE)

cat(sprintf("Mean Age: %.2f, Median Age: %.2f", mean_age, median_age))
```

```{r}
library(ggplot2)

# Histogram plot of Age with color based on Gender
SurveyData$Gender <- factor(SurveyData$Gender, levels = c("Male", "Female"))

ggplot(data = SurveyData, aes(x = Age, fill = Gender)) +
  geom_histogram(
    position = "stack",
    binwidth = 3,
    colour = "grey"
  ) +
  labs(title = "Number of Male and Female by Age") +
  theme_minimal()
```


## Income Category Variable Created ##
```{r}
# Define labels for the income categories
breaks <- c(0, 200000, 600000, 1200000, 2000000, Inf)
labels <-
  c(
    "Low Income",
    "Lower-Middle Income",
    "Middle-Middle Income",
    "Upper-Middle Income",
    "High Income"
  )

# Create the IncomeCategory variable
SurveyData$IncomeCategory <-
  cut(
    SurveyData$HouseholdIncome,
    breaks = breaks,
    labels = labels,
    include.lowest = TRUE
  )

# add var to the dataframe
write.csv(SurveyData, "SurveyData.csv", row.names = FALSE)
```


## Variable for TotalTechConcern Created ##
```{r}
library(dplyr)

# TotalTechConcern calculated by summing all columns that start with 'Risk'
SurveyData <- SurveyData %>%
  mutate(TotalTechConcern = rowSums(select(., starts_with("Risk"))))

# add var to the dataframe
write.csv(SurveyData, "SurveyData.csv", row.names = FALSE)
```

## Variable for TotalTechBenefit Created ##
```{r}
library(dplyr)

# Calculate the TotalTechBenefit by summing all columns that start with 'Benefit'
SurveyData <- SurveyData %>%
  mutate(TotalTechBenefit = rowSums(select(., starts_with("Benefit")), na.rm = TRUE))

# add var to the dataframe
write.csv(SurveyData, "SurveyData.csv", row.names = FALSE)
```


## Plots that look at the questions ##

Plotting the correlation between Age and TotalTechConcern
    Relevant Question: Does age-group affect appraisals of technology?

```{r}
library(ggplot2)

ggplot(data = SurveyData, aes(x = Age, y = TotalTechConcern)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", color = "blue") +
  labs(
    x = "Age", y = "Total Tech Concern",
    title = "Correlation between Age and Total Tech Concern"
  ) +
  theme_minimal()

# Calculate and print the correlation coefficient
correlation <-
  cor.test(SurveyData$Age, SurveyData$TotalTechConcern, use = "complete.obs") # use handles missing values
print(correlation)
```
 


## Plotting the relationship between IncomeCategory and TotalTechConcern ##
     Relevant Question: Whether income-level negatively affects an individuals appraisals of technology (lower = more neg)
    
```{r}
library(ggplot2)

ggplot(data = SurveyData, aes(x = IncomeCategory, y = TotalTechConcern)) +
  geom_boxplot() +
  labs(
    x = "Income Category", y = "Total Tech Concern",
    title = "Total Tech Concern across Income Categories"
  ) +
  theme_minimal()

# ANOVA to test if differences in TotalTechConcern are statistically significant across IncomeCategory
anova_result <- aov(TotalTechConcern ~ IncomeCategory, data = SurveyData)
summary(anova_result)
```
# Results & Follow up 

P > 0.05 (No significant effect) for both observations. Therefore, income-level and age-group does not correlate with an individuals appraisals of technology. 

Problem of Gender Imbalance: To counter the overrepresentation of male participants, future iterations of the survey could aim for a more balanced gender representation.

Diversifying Geographic Representation: The data is gathered from participants residing in India. As a result, generalizability is called into question. Conducting the survey in different countries or regions would provide comparative insights and reveal whether the trends observed are specific to India or more universally applicable. 

Fit for Purpose: The data aligns well with the specific research questions I have, however, the sample representativeness is a weak point of the data. generalizability is not possible with this data, therefore, it is not fit for purpose. Regardless, the data provides a good starting point.  


# Reviewer Notes: 

*Overall, the code runs smoothly and it makes sense. I certianly think your current code addresses your research questions. Additionall, I think the research questions are very interesting! However, there are some issues, some of which are big. Notably, do not overwrite your original data! You can add the variables you create to your loaded data without overwriting the data you have saved in your folder. Always preserve the original data. Instead, if you would like to have a data set with these variables, write out a version of the data (with a different name) at the end of your code (before your analysis plots) that writes out this data. Please see my uploaded code review spreadsheet for further comments.*

*To answer the specific questions:*

*1. It is clear what steps I need to do to run the script.*

*2.When I take these steps, the script runs smoothly with no errors.*

*3. The html file is largely reproducible. I get all of the same output and results. However, in order to get the .RMD file to knit, I have to add eval = FALSE to line 15. So, I am unsure how the author got the code to knit without this.*

*4. It is largely easy to read the scirpt and understand what the author is doing and why. The only point of contention is why the author overwrites the original data with the newly created variables (see above note and code review spreadsheet for further explanation). I understand what the code is doing but not why the author decides to do this.*


